### 2020-01-23 README.Rmd
add ShinyProxy, update list in README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Rockerverse

<!-- badges: start -->
[![Launch RStudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/nuest/rockerverse-paper/master?urlpath=rstudio)
[![Build Status](https://travis-ci.org/nuest/rockerverse-paper.svg?branch=master)](https://travis-ci.org/nuest/rockerverse-paper)
<!-- badges: end -->

Draft for an article of everything related to containers and R, originally based on the blog post http://bit.ly/docker-r.

_Want to contribute?_ See https://github.com/nuest/rockerverse-paper/issues/3.

**Current draft: https://nuest.github.io/rockerverse-paper/master.pdf**

```{r hexSticker, message=TRUE, warning=FALSE, cache=TRUE, include=FALSE}
if(!file.exists("rockerverse.png")) {
  library("magick")
  library("here")

  # https://commons.wikimedia.org/wiki/File:NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg
  universe <- magick::image_read("https://upload.wikimedia.org/wikipedia/commons/6/69/NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg")
  magick::image_write(universe, here("universe.png"))
  # rockerLogoMask <-
  magick::image_read("https://avatars2.githubusercontent.com/u/9100160?s=200&v=4") %>%
    magick::image_scale("2100x2100") %>%
    magick::image_channel("Matte") %>%
    image_negate() %>%
    magick::image_border("black", "150x0") %>%
    magick::image_write(here("rocker_mask.png"))

  # does not work: image_composite(universe, rockerLogoMask, operator = "CopyOpacity", compose_args = "-composite -alpha Off")
  system2("convert", c(here("universe.png"), here("rocker_mask.png"),
    "-alpha Off",
    "-compose CopyOpacity",
    "-composite",
    here("whaleverse.png")))

  library("hexSticker")
  hexSticker::sticker(here("whaleverse.png"), package = "Rockerverse",
          p_size = 20, s_x = 0.97, s_y = 0.95, s_width = 1,
          h_fill = "#000000", h_color = "#ffffff",
          url = "github.com/nuest/rockerverse-paper",
          u_color = "#ffffff", u_size = 4,
          filename = here("rockerverse.png"))

  file.remove(c(here("universe.png"), here("whaleverse.png"), here("rocker_mask.png")))
}
```

![Rockerverse hex sticker](`r here::here("rockerverse.png")`)

## Render the manuscript with Docker

**Local build**

```bash
docker build --tag rockerverse-paper --file .binder/Dockerfile .
docker run -i -v $(pwd):/rockerverse --user $UID rockerverse-paper Rscript -e 'setwd("/rockerverse"); rmarkdown::render("manuscript.Rmd")'
```

**Image from Docker Hub**

```bash
docker run -i -v $(pwd):/rockerverse --user $UID nuest/rockerverse-paper Rscript -e 'setwd("/rockerverse"); rmarkdown::render("manuscript.Rmd")'
```

## Not included projects and packages

- `rize`, because discontinued, see #35
- `docker` ([https://bhaskarvk.github.io/docker](https://bhaskarvk.github.io/docker) has been removed from CRAN, seems unmaintained by authors
- https://github.com/svlentink/dockerfiles/blob/master/svlentink/r-base-alpine/Dockerfile because the Alpine image is quite a trivial one
- https://github.com/jlisic/R-docker-centos because CentOS images are also provided by R-hub
- `sevenbridges` includes several Dockerfiles and documentation about using images, and also seems to use Docker in the built workflows but via other tools, not directly, https://sbg.github.io/sevenbridges-r/, https://github.com/sbg/sevenbridges-r/search?p=4&q=docker&unscoped_q=docker
- https://github.com/rocker-jp because it seems to have stalled soon after inception
- https://github.com/thomasp85/fiery because no examples/docs for Docker (though probably trivial anyway)
- [RStudio Connect](https://rstudio.com/products/connect/) because while it does use a similar mechanism to seperate processes, it does _not_ use Docker and the comparison would require a lot of background explanation, and not all details are available because the software is proprietary