### 2020-01-31 README.Rmd
use more suitable font for sticker

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Rockerverse

<!-- badges: start -->
[![Launch RStudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/nuest/rockerverse-paper/master?urlpath=rstudio)
[![Build Status](https://travis-ci.org/nuest/rockerverse-paper.svg?branch=master)](https://travis-ci.org/nuest/rockerverse-paper)
<!-- badges: end -->

Draft for an article of everything related to containers and R, originally based on the blog post http://bit.ly/docker-r.

_Want to contribute?_ See https://github.com/nuest/rockerverse-paper/issues/3.

**Current draft: https://nuest.github.io/rockerverse-paper/master.pdf**

```{r hexSticker, message=TRUE, warning=FALSE, cache=TRUE, include=FALSE}
library("magick")
library("here")

# https://commons.wikimedia.org/wiki/File:NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg
if(!file.exists("universe.png")) {
  universe <- magick::image_read("https://upload.wikimedia.org/wikipedia/commons/6/69/NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg")
  magick::image_write(universe, here("universe.png"))
}

magick::image_read("https://avatars2.githubusercontent.com/u/9100160?s=200&v=4") %>%
  magick::image_scale("2100x2100") %>%
  magick::image_channel("Matte") %>%
  image_negate() %>%
  magick::image_border("black", "150x0") %>%
  magick::image_write(here("rocker_mask.png"))

# does not work: image_composite(universe, rockerLogoMask, operator = "CopyOpacity", compose_args = "-composite -alpha Off")
system2("convert", c(here("universe.png"), here("rocker_mask.png"),
  "-alpha Off",
  "-compose CopyOpacity",
  "-composite",
  here("whaleverse.png")))

library("sysfonts")
if(!file.exists(here::here("FinalFrontierOldStyle-8Pg.ttf"))) {
  stop("Download font Final Frontier Old Style, e.g. from https://www.fontspace.com/allen-r-walden/final-frontier-old-style and put the file in the repo root.")
}
sysfonts::font_add("finalfrontier", here::here("FinalFrontierOldStyle-8Pg.ttf"))

library("hexSticker")
hexSticker::sticker(here::here("whaleverse.png"), package = "ROCKERVERSE",
        p_family = "finalfrontier",
        p_size = 29, p_y = 1.3,
        s_x = 0.97, s_y = 0.95, s_width = 1,
        h_fill = "#000000", h_color = "#ffffff",
        url = "github.com/nuest/rockerverse-paper",
        u_color = "#ffffff", u_size = 4,
        filename = here::here("rockerverse.png"))

#file.remove(c(here("universe.png"), here("whaleverse.png"), here("rocker_mask.png")))
```

![Rockerverse hex sticker](`r here::here("rockerverse.png")`)

Hex sticker uses the font [Final Frontier Old Style](https://www.fontspace.com/allen-r-walden/final-frontier-old-style) and the [Hubble Ultra-Deep Field](https://en.wikipedia.org/wiki/Hubble_Ultra-Deep_Field) combined with the Docker logo as background.

## Render the manuscript with Docker

**Local build**

```bash
docker build --tag rockerverse-paper --file .binder/Dockerfile .
docker run -i -v $(pwd):/rockerverse --user $UID rockerverse-paper Rscript -e 'setwd("/rockerverse"); rmarkdown::render("manuscript.Rmd")'
```

**Image from Docker Hub**

```bash
docker run -i -v $(pwd):/rockerverse --user $UID nuest/rockerverse-paper Rscript -e 'setwd("/rockerverse"); rmarkdown::render("manuscript.Rmd")'
```

## Not included projects and packages

- `rize`, because discontinued, see #35
- `docker` ([https://bhaskarvk.github.io/docker](https://bhaskarvk.github.io/docker) has been removed from CRAN, seems unmaintained by authors
- https://github.com/svlentink/dockerfiles/blob/master/svlentink/r-base-alpine/Dockerfile because the Alpine image is quite a trivial one
- https://github.com/jlisic/R-docker-centos because CentOS images are also provided by R-hub
- `sevenbridges` includes several Dockerfiles and documentation about using images, and also seems to use Docker in the built workflows but via other tools, not directly, https://sbg.github.io/sevenbridges-r/, https://github.com/sbg/sevenbridges-r/search?p=4&q=docker&unscoped_q=docker
- https://github.com/rocker-jp because it seems to have stalled soon after inception
- https://github.com/thomasp85/fiery because no examples/docs for Docker (though probably trivial anyway)
- [RStudio Connect](https://rstudio.com/products/connect/) because while it does use a similar mechanism to seperate processes, it does _not_ use Docker and the comparison would require a lot of background explanation, and not all details are available because the software is proprietary
- https://github.com/ecohealthalliance/reservoir is an invididual specific data science image container for a group - add if it fits
- https://itsalocke.com/blog/building-an-r-training-environment/ (one single container for many users in teaching, uses rocker/tidyverse)