### 2020-03-03 README.Rmd
move helper functions to README, add rockerverse.R
preparing for submission, see #26

also add missing used packages to DESCRIPTION

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Rockerverse

<!-- badges: start -->
[![Launch RStudio Binder](http://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/nuest/rockerverse-paper/master?urlpath=rstudio)
[![Build Status](https://travis-ci.org/nuest/rockerverse-paper.svg?branch=master)](https://travis-ci.org/nuest/rockerverse-paper)
<!-- badges: end -->

Article about everything related to containers and R, originally based on the blog post http://bit.ly/docker-r and now **available as a preprint on arXiv.org: https://arxiv.org/abs/2001.10641**.

_Want to contribute?_ See https://github.com/nuest/rockerverse-paper/issues/3.

**Current draft: https://nuest.github.io/rockerverse-paper/master.pdf**

```{r hexSticker, message=TRUE, warning=FALSE, cache=TRUE, include=FALSE}
library("magick")
library("here")

# https://commons.wikimedia.org/wiki/File:NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg
if(!file.exists("universe.png")) {
  universe <- magick::image_read("https://upload.wikimedia.org/wikipedia/commons/6/69/NASA-HS201427a-HubbleUltraDeepField2014-20140603.jpg")
  magick::image_write(universe, here("universe.png"))
}

magick::image_read("https://avatars2.githubusercontent.com/u/9100160?s=200&v=4") %>%
  magick::image_scale("2100x2100") %>%
  magick::image_channel("Matte") %>%
  image_negate() %>%
  magick::image_border("black", "150x0") %>%
  magick::image_write(here("rocker_mask.png"))

# does not work: image_composite(universe, rockerLogoMask, operator = "CopyOpacity", compose_args = "-composite -alpha Off")
system2("convert", c(here("universe.png"), here("rocker_mask.png"),
  "-alpha Off",
  "-compose CopyOpacity",
  "-composite",
  here("whaleverse.png")))

library("sysfonts")
if(!file.exists(here::here("FinalFrontierOldStyle-8Pg.ttf"))) {
  stop("Download font Final Frontier Old Style, e.g. from https://www.fontspace.com/allen-r-walden/final-frontier-old-style and put the file in the repo root.")
}
sysfonts::font_add("finalfrontier", here::here("FinalFrontierOldStyle-8Pg.ttf"))

library("hexSticker")
factor <- 1 # for print: 4
hexSticker::sticker(here::here("whaleverse.png"), package = "ROCKERVERSE",
        p_family = "finalfrontier",
        p_size = 29 * factor, p_y = 1.3,
        s_x = 0.97, s_y = 0.95, s_width = 1,
        h_fill = "#000000", h_color = "#ffffff",
        url = "github.com/nuest/rockerverse-paper",
        u_color = "#ffffff", u_size = 4 * factor,
        filename = here::here(ifelse(
          factor == 1,
          "rockerverse.png",
          paste0("rockerverse_", 300 * factor, "dpi.png"))),
        dpi = 300 * factor)

#file.remove(c(here("universe.png"), here("whaleverse.png"), here("rocker_mask.png")))
```

![Rockerverse hex sticker](rockerverse.png)

The hex sticker uses the font [Final Frontier Old Style](https://www.fontspace.com/allen-r-walden/final-frontier-old-style) and the [Hubble Ultra-Deep Field](https://en.wikipedia.org/wiki/Hubble_Ultra-Deep_Field) combined with the Docker logo as background.

## Render the manuscript with Docker

**Local build**

```bash
docker build --tag rockerverse-paper --file .binder/Dockerfile .
docker run -i -v $(pwd):/rockerverse --user $UID rockerverse-paper Rscript -e 'setwd("/rockerverse"); rmarkdown::render("rockerverse.Rmd")'
```

**Image from Docker Hub**

```bash
docker run -i -v $(pwd):/rockerverse --user $UID nuest/rockerverse-paper Rscript -e 'setwd("/rockerverse"); rmarkdown::render("rockerverse.Rmd")'
```

## Not included projects and packages

- `rize`, because discontinued, see #35
- `docker` ([https://bhaskarvk.github.io/docker](https://bhaskarvk.github.io/docker) has been removed from CRAN, seems unmaintained by authors
- https://github.com/svlentink/dockerfiles/blob/master/svlentink/r-base-alpine/Dockerfile because the Alpine image is quite a trivial one
- https://github.com/jlisic/R-docker-centos because CentOS images are also provided by R-hub
- `sevenbridges` includes several Dockerfiles and documentation about using images, and also seems to use Docker in the built workflows but via other tools, not directly, https://sbg.github.io/sevenbridges-r/, https://github.com/sbg/sevenbridges-r/search?p=4&q=docker&unscoped_q=docker
- https://github.com/rocker-jp because it seems to have stalled soon after inception
- https://github.com/thomasp85/fiery because no examples/docs for Docker (though probably trivial anyway)
- [RStudio Connect](https://rstudio.com/products/connect/) because while it does use a similar mechanism to seperate processes, it does _not_ use Docker and the comparison would require a lot of background explanation, and not all details are available because the software is proprietary
- https://github.com/ecohealthalliance/reservoir is an invididual specific data science image container for a group - add if it fits
- https://itsalocke.com/blog/building-an-r-training-environment/ (one single container for many users in teaching, uses rocker/tidyverse)

## Helpers

The [README.Rmd](README.Rmd) file's source includes some useful little code snippets, e.g. count words or get a list of all contributor's GitHub handles.

```{r word_stats, eval=FALSE, include=FALSE}
# devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
wordcountaddin::readability(here::here("rockerverse.Rmd"))
wordcountaddin::text_stats(here::here("rockerverse.Rmd"))
```

```{r preprint_pdf, eval=FALSE, include=FALSE}
# run this chunk, get the PDF, then revert changes in touched files
library("rmarkdown")
library("rticles")
unlink('manuscript_cache', recursive = TRUE)

# function is a copy of rticles::rjournal_article() but...
# - using RJWrapper_preprint_template.tex (which includes header.tex commands)
# - directly referencing unexported functions from rticles
# - direclty setting the keep_tex option
rjournal_style_preprint <- function (..., citation_package = "natbib") {
    rmarkdown::pandoc_available("2.2", TRUE)
    base <- rticles:::pdf_document_format("rjournal_article", highlight = NULL,
        citation_package = citation_package, ...)
    base$pandoc$to <- "latex"
    base$pandoc$ext <- ".tex"
    base$pandoc$keep_tex <- TRUE # added
    base$post_processor <- function(metadata, utf8_input, output_file,
        clean, verbose) {
      filename <- basename(output_file)
      if (filename != (filename2 <- gsub("_", "-", filename))) {
          file.rename(filename, filename2)
          filename <- filename2
      }
      m <- list(filename = xfun::sans_ext(filename))
      h <- rticles:::get_list_element(metadata, c("output", "rticles::rjournal_article",
          "includes", "in_header"))
      h <- c(h, if (length(preamble <- unlist(metadata[c("preamble",
          "header-includes")]))) {
          f <- tempfile(fileext = ".tex")
          on.exit(unlink(f), add = TRUE)
          xfun::write_utf8(preamble, f)
          f
      })
      #t <- find_resource("rjournal_article", "RJwrapper.tex")
      t <- here::here("RJwrapper_preprint_template.tex")
      rticles:::template_pandoc(m, t, "RJwrapper.tex", h, verbose)
      tinytex::latexmk("RJwrapper.tex", base$pandoc$latex_engine,
          clean = clean)
    }
    base$knitr$opts_chunk$comment <- "#>"
    hilight_source <- rticles:::knitr_fun("hilight_source")
    hook_chunk = function(x, options) {
        if (rticles:::output_asis(x, options))
            return(x)
        paste0("```{=latex}\n\\begin{Schunk}\n", x, "\\end{Schunk}\n```")
    }
    hook_input <- function(x, options) paste(c("\\begin{Sinput}",
        hilight_source(x, "sweave", options), "\\end{Sinput}",
        ""), collapse = "\n")
    hook_output <- function(x, options) paste("\\begin{Soutput}\n",
        x, "\\end{Soutput}\n", sep = "")
    base$knitr$knit_hooks <- rticles:::merge_list(base$knitr$knit_hooks,
        list(chunk = hook_chunk, source = hook_input, output = hook_output,
            message = hook_output, warning = hook_output, plot = knitr::hook_plot_tex))
    base
}

rmarkdown::render("rockerverse.Rmd", output_format = rjournal_style_preprint())
file.rename("RJwrapper.pdf", "rockerverse-preprint.pdf") # file for own archival

# For submission to arXiv:
# 1. Copy the following files into a new folder:
# - figures/*
# - rockerverse.tex
# - RJournal.sty
# - RJreferences.bib
# - RJwrapper.tex
# 2. Run pdfLaTeX directly on RJwrapper.tex (works with RStudio)
# 3. Zip the following files and submit to arXiv:
# - /figures
# - rockerverse.tex
# - RJournal.sty
# - RJwrapper.bbl
# - RJwrapper.tex
```

```{r collab_emails, eval=FALSE, include=FALSE}
front_matter <- rmarkdown::yaml_front_matter("rockerverse.Rmd")
emails <- sapply(X = front_matter$author, FUN = function(a) { a$email })
cat(emails, sep = "\n")
```

```{bash collab_handles, eval=FALSE, include=FALSE}
cat rockerverse.Rmd | grep '    # https://github.com/' | sed 's|    # https://github.com/|@|'
```